3D Regression Heat Map Analysis of Population Study Data

### 2015-06-25 draft.tex
removed color codes

Testing features for associations with diseases using regression models is one of the most important epidemiological tools.
%%
Using regression analysis to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
%%
Due to the amount of data and only limited overview visualizations, possible correlations may be missed.
%%
Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target feature.
%%
They mostly highlight correlations between features, which are known to the domain expert (e.g., correlation between body size and spine shape).
%%
We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis w.r.t. a specific disease or hypothesis.
%%
For this purpose we provide template regression formulas, which are applied to all potential feature combinations.
%%
Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
%%
Difference views between regression formulas allow to assess the influences of individual features on the process.
%%
Our contributions are:

\design{
\begin{itemize}
	%%
  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis generation w.r.t. a target feature using regression models.
	%%
  %\item Applying domain knowledge in the analysis using freely adjustable regression formulas.
	\item Incorporation of prior domain knowledge by using freely adjustable regression formulas.
	%%
  %\item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details-on-demand to assess the model quality.
	\item Metrics selection for analyzing regression models and for details-on-demand representations.
	%%
  %\item An openly available web-based system, which incorporates open formats to be applicable for data of other application domains.
	\item An open-source web-application that can be used with data of different application domains.
	%%
\end{itemize}
}
% \begin{itemize}
% 	%%
% 	\item An overview visualization technique representing feature interactions using target features.
% 	%%
% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details-on-demand techniques for in-depth investigations of feature relationships.
% 	%%
% 	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
% 	%%
% 	\item The analysis of confounding features by providing comparison views between different formula results.
% 	%%
% 	\item The expansion and adaption of our method to any data with our open web-based system.
% 	%%
% \end{itemize}
%%
% \noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
% %%
% They extend 2D scatter plots to display detailed information about a regression model.
% %%
% In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
% %%
\section{Epidemiological Background} \label{sec:Background}
%%
This section covers the epidemiological workflow, epidemiological data in general and the SHIP in particular.
%%
\subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
%%
Epidemiological research is performed by experts from different academic disciplines, such as epidemiologists, physicians, statisticians, and medical computer scientists focusing on biometrics and image segmentation.
%%
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
%%
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{enumerate}
	\item Clinicians/Epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.
	%\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features w.r.t the investigated disease.
\end{enumerate}
%%
Relative risks can be determined if a statistical resilient association of features with a condition is extracted.
%%
Derived absolute risks indicate the per-subject chance of developing the disease.
%%
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
%%
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow.
%%
Hence, images are mostly used for the communication of results, rather than for providing insight into the data.
%%

@@

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data and are the basis of this work.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
\add{
%%
The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model by estimating its information loss of modeling the underlying data and is suited for comparing models \cite{Akaike}.
%%
A small AIC value indicates a higher quality model.
%%
}
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value ranges between $[0, 1]$, where 1 encodes a perfect fit.
%%

\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
%%
$Dependent \sim Independent_1 + ... + Independent_n$.
% \begin{equation}
% Dependent \sim Independent_1 + ... + Independent_n
% \label{eq:RegressionNotation}
% \end{equation}
%%
An example of a regression formula would be $KidneyDisorder~\sim~Smoking~+~Obesity$.
The most commonly used regression operators comprise:
%%
\begin{itemize}
	\item $+,-$ inclusion/exclusion of the variable (e.g. $x\pm y$),
	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
	\item $*$ inclusion of the variables as well as their interactions (e.g. $x*y$)
	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
\end{itemize}
%%
\magenta{
Due to the different meaning of operators in regression formulas, feature transformations are not available within this notation.
}
%%
The class of the target feature restricts the regression type.
%%
Different regression types are available, we focus on the following for describing linear relationships:
%%
\paragraph{Linear Regression for Continuous Target.}
%%
The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
%%
The \emph{dependent} variable has to be of \emph{continuous type}.
%%
\add{
%%
Linear regression models can also be described using the \emph{adjusted} $R^2$, which considers the number of dependent features.
%%
It yields lower values, when features with little entropy w.r.t. the target feature are included.
%Additional metrics describing linear regression models comprise of the \emph{adjusted} $R^2$, which also adjusts for the number of dependent features and yields lower values, when features with little entropy towards the target feature are included.
%%
%The $f$-$statistic$ determines how much improvement the independent variables will add to the model compared to being not considered at all.
The $f$-$statistic$ measures the improvement of the model if independent variables were added. 
%%
}

\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
%%
The target is described by fitting a logistic function.
%%
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
%%
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$} \cite{Nagelkerke}, which mimics the behavior of the $R^2$.
%%
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
%%
Comparisons have to be handled with care.
%%
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
Tukey already stated in 1977 that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
%%
He emphasized the need to use data to \emph{derive} hypotheses, which can then be tested again.
%%
In this section we present prior and related work trying to achieve this goal.
%%
%The work of M\"{u}hlbacher et al.~\cite{Muehlbacher}, Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
\paragraph{Parameter space analysis using regression models.}
%%
\magenta{
Sedlmair et al.~\cite{Sedlmair} presented a conceptual taxonomy of parameter space analysis.
%%
%It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
%%
%Using the taxonomy, approaches can be classified to find works with similar problems and derive new design ideas and evaluate ideas.
%The taxonomy allows to classify approaches to find works with similar problems to derive new design ideas and evaluate ideas.
%%
Based on their input parameter taxonomy, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
%%
We solve a \emph{fitting} task by aiming to find models well suited for describing the input data.
}

\magenta{
%% A Partition-based Framework for Building and Validating Regression Models
The approach of M\"{u}hlbacher et al.~\cite{Muehlbacher} is closest to ours.
%%
They provided a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
%%
%Correlations between features with the target yield a ordered matrix-plot, where feature combinations are used to depict models of interest.
%A view for single-relationships with the target and single features as well as a matrix-plot for feature combinations are used to depict models of interest.
%%
%The visualization allows to assess different model complexities.
%%
Existing regression models can be validated and compared using 3D views and 2D slice views.
%%
%We employ the same steps by preprocessing features using correlations with the target and providing an overview over combinations.
%%
M\"{u}hlbacher et al. focused on a smaller number of features, which can be assessed in more detail, yielding a plot matrix view, while we cover more features by abstracting the models.
%%
%We also take categorical targets and regression formula inputs into account.
%%
}

%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Similarly, Piringer et al.~\cite{Piringer} proposed methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to assess the pairwise influence of independent features w.r.t. the target feature using a plot matrix displaying models as contours.
%%
%They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
%%
Linked views of model deviations allow to select outliers.
%%
This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
%%
% The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%
% We analyze a large amount of simpler models in terms of \emph{different} features.
% %%
% They also focus on metric features, while we process categorical data as well.
%%
\magenta{
%%
Guo et al.~\cite{Guo} presented multi-space visualizations to find linear relationships in the data with focus on extracting groups of best fit.
%%
The data space is visualized using a scatter plot matrix.
%%
Linear models are calculated by defining dependent and independent features.
%%
The model view allows for assessing different models by color coding distances to the line-of-fit.
%%
The model-parameters can then be fine-tuned using line graphs, histograms and model projections.
%%
}
%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
%We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
%%
Chan et al.~\cite{Chan} propose the `Regression Cube', an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to Piringer et al.~\cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
%In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
%We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%
Piringer et al.~\cite{Piringer}, Guo et al.~\cite{Guo} and Chan et al.~\cite{Chan} focus on finding and tuning a model for a specific relationship, while we search for models w.r.t. all combinations in the data set.
%%
Instead of analyzing one complex model in detail, we process a large amount of models in terms of \emph{different} features.
%We difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%

@@

%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
Steenwijk et al.~\cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
%%
They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
%%
They incorporate linked views using scatter plots, bar charts, parallel coordinates as well as time plots to display and brush the data.
%%
%Medical image data can be incorporated to investigate potentially interesting individuals.
%%
%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
%%
%They are therefore suitable for \emph{small} epidemiological data sets.
%%
\magenta{
Maries et al.~\cite{GRACE} proposed \texttt{GRACE}, a framework for visually exploring correlations of features with spatial and non-spatial geriatric data.
%%
Sparse Partial Least Squares Regression and Tikhonov Regularization are used to produce predictor subsets and quantify correlation strengths.
%%
A multiple view system links spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams, correlations are plotted using p-values in the other views.
%%
%Dendograms are used to select different brain regions.
Instead of applying several linked views, we focus on one view of abstracted data.
}

Generalized Pairs plots (\texttt{GPLOM'S}) extend the concept of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
Dai et al.~\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
%%
A \emph{concept map} summarizes features related to a specified disease.
%%
Time-dependent epidemiological data are visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
%%
\magenta{
We incorporate the idea of the arrangement of feature combinations as matrix as well as assessing confounding features.
%%
}
%%
%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Statistical analysis.}
%%
%Statistical analysis is an inherent part of Visual Analysis.
%%
%Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high dimensional-data.
%%
Their refer to their research agenda in our design.
%%
Namely, we apply aspects of perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high dimensional data into subspaces, which are then described by their most reliable linear regression model.
%%
%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on an algebraic representation for efficient regression-model calculation to find the best fit for a subspace.
%%
%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

@@

\magenta{

@@

\design{

@@

As described in Section~\ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing relationships in epidemiological data.
%%
A regression model is based on expert knowledge.
%%
There is no rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
% \design{
% %%
% Key requirements based on the epidemiological application domain on the 3D Regression Heat Map are:
% %%
% \begin{itemize}
% 	\item Show relationships of features w.r.t. their explanatory power towards the target feature (e.g., a disease or unhealthy condition).
% 	\item Allow for modeling assumptions regarding the target feature.
% 	\item The tool has to be easily accessible and simultaneously secure the data its processing.
% \end{itemize}
% %%
% }
\subsection{Regression Heat Map Description Using Regression Formula Notation}
%%
\design{Expert knowledge modeling} is carried out using \emph{regression formulas}.
%Expert knowledge is integrated into the analysis using regression formulas.
%%
The formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target (recall Sect.~\ref{sec:RegressionAnalysis}).
%%

@@

The number of features calculated by the CFS algorithm is dependent on the information entropy in the data.
%%
In our epidemiological data, we usually observed a number of 10 to 30 features.
%%
The number of selected features using the CFS algorithm reflects their information entropy on the target.
%%
A large list of features is an expression of low correlation to the target feature.
%%
\magenta{
The tradeoff involved using the CFS algorithm is the potential removal of interesting features for the domain expert.
%%
This problem is discussed in the next subsection as part of the 3D representation of the regression results.
}
%%

@@

\paragraph{3D Regression Heat Map comparison.}
\design{
%%
Comparisons were introduced later in the project.
%%
Epidemiologists with focus on statistics pointed out that comparing outcomes of different formulas is suitable for removing the effect of possible confounding features.
%%
}
%%
\emph{3D Regression Heat Maps} can be compared by creating difference views.
%%
One formula acts as reference.
%%
The absolute difference in the regression metric values with the second formula is calculated. %yielding the difference containing only the differences between the two formulas.
%%
For example, it can be utilized for comparing the influence of a single feature on the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$).

>>>>

Testing features for associations with diseases using regression models is one of the most important epidemiological tools.
Using regression analysis to assess the statistical resilience of a hypothesis rarely involves \emph{more than three features} due to the higher dimensional problem and the required subject count.
Due to the amount of data and only limited overview visualizations, possible correlations may be missed.
Explorative analyses and overview visualizations of the data set as presented in prior work \cite{Klemm2014VIS} are not tailored to a specific target feature.
They mostly highlight correlations between features, which are known to the domain expert (e.g., correlation between body size and spine shape).
We incorporate the regression analysis, which is familiar to the domain experts, into overview visualizations to support a hypothesis-free analysis or an analysis w.r.t. a specific disease or hypothesis.
For this purpose we provide template regression formulas, which are applied to all potential feature combinations.
Since the notation is familiar to epidemiologists, they can rapidly include their domain knowledge into the analysis process.
Difference views between regression formulas allow to assess the influences of individual features on the process.
Our contributions are:
\begin{itemize}
  \item An overview visualization design based on feedback of epidemiological domain experts to support hypothesis generation w.r.t. a target feature using regression models.
  %\item Applying domain knowledge in the analysis using freely adjustable regression formulas.
	\item Incorporation of prior domain knowledge by using freely adjustable regression formulas.
  %\item Selecting suitable metrics for representing and comparing regression models, as well as incorporating additional metrics for details-on-demand to assess the model quality.
	\item Metrics selection for analyzing regression models and for details-on-demand representations.
  %\item An openly available web-based system, which incorporates open formats to be applicable for data of other application domains.
	\item An open-source web-application that can be used with data of different application domains.
\end{itemize}
% \begin{itemize}
% 	\item An overview visualization technique representing feature interactions using target features.
% 	\item Visualization techniques, which incorporate overview visualizations of all regression analyses at once as well as details-on-demand techniques for in-depth investigations of feature relationships.
% 	\item Freely adjustable regression formulas to provide a simple, yet powerful way to adjust the regression analysis to specific hypotheses about the data.
% 	\item The analysis of confounding features by providing comparison views between different formula results.
% 	\item The expansion and adaption of our method to any data with our open web-based system.
% \end{itemize}
% \noindent The term `Regression Cube' has been used for the detailed analysis of one or a small number of regression models \cite{Ahmadi, Chan}.
% They extend 2D scatter plots to display detailed information about a regression model.
% In behalf of that, we declare our visualization technique for comparing and analyzing a large number of different regression models with different input features as \emph{Regression Cube}.
\section{Epidemiological Background} \label{sec:Background}
This section covers the epidemiological workflow, epidemiological data in general and the SHIP in particular.
\subsection{Epidemiological Workflow} \label{sec:EpidemiologicalWorkflow}
Epidemiological research is performed by experts from different academic disciplines, such as epidemiologists, physicians, statisticians, and medical computer scientists focusing on biometrics and image segmentation.
Their goal is to derive disease-specific risk factors by assessing epidemiological features with statistical methods.
As described by Thew et al.~\cite{Thew2009} the epidemiological workflow is divided into these different steps:
\begin{enumerate}
	\item Clinicians/Epidemiologists derive hypotheses via clinical observations, experimental studies or literature research.
	%\item Clinicians make observations in the daily practice, which are translated into hypotheses.
	\item Epidemiologists compile a list of features depicting the hypothesis and include confounding features.
	\item Statisticians assess the association of the derived features w.r.t the investigated disease.
\end{enumerate}
Relative risks can be determined if a statistical resilient association of features with a condition is extracted.
Derived absolute risks indicate the per-subject chance of developing the disease.
Reproducibility of results is an epidemiological key requirement and guides all analyses steps.
Statistical programs, such as \texttt{SPSS}, are used to analyze the data regarding the classical sequential epidemiological workflow.
Hence, images are mostly used for the communication of results, rather than for providing insight into the data.

@@

\subsection{Regression Analysis} \label{sec:RegressionAnalysis}
%%
Regression analysis is the most important statistical tool when analyzing epidemiological data and are the basis of this work.
%%
A regression analysis assesses the influence of one or more (\emph{independent}) features to one target (\emph{dependent}) feature.
%%
The regression model yields a function describing the target feature by weighting the independent features.
%%
Different metrics, such as the weightings itself and associated $p$ values, describe the resulting function (the \emph{model}).
%%
The \emph{Akaike Information Criterion} (AIC) metric estimates the quality of a model by estimating its information loss of modeling the underlying data and is suited for comparing models \cite{Akaike}.
%%
A small AIC value indicates a higher quality model.
%%
$R^2$ values describe the quality of the fit; in other words how well the dependent features describe the target feature.
%%
The value ranges between $[0, 1]$, where 1 encodes a perfect fit.
%%

\paragraph{Regression Analysis Notation.} Regression formulas are usually denoted as follows:
$Dependent \sim Independent_1 + ... + Independent_n$.
% \begin{equation}
% Dependent \sim Independent_1 + ... + Independent_n
% \label{eq:RegressionNotation}
% \end{equation}
An example of a regression formula would be $KidneyDisorder~\sim~Smoking~+~Obesity$.
The most commonly used regression operators comprise:
\begin{itemize}
	\item $+,-$ inclusion/exclusion of the variable (e.g. $x\pm y$),
	\item $:$ inclusion of interactions between the variables (e.g. $x:y$),
	\item $*$ inclusion of the variables as well as their interactions (e.g. $x*y$)
	\item $|$ (conditioning) inclusion of variable x, given y (e.g. $x|y$)
\end{itemize}
Due to the different meaning of operators in regression formulas, feature transformations are not available within this notation.
The class of the target feature restricts the regression type.
Different regression types are available, we focus on the following for describing linear relationships:
\paragraph{Linear Regression for Continuous Target.}
The basic type is the linear regression, creating a linear map from the space comprising the \emph{independent} features to the \emph{dependent} features.
The \emph{dependent} variable has to be of \emph{continuous type}.
Linear regression models can also be described using the \emph{adjusted} $R^2$, which considers the number of dependent features.
It yields lower values, when features with little entropy w.r.t. the target feature are included.
%Additional metrics describing linear regression models comprise of the \emph{adjusted} $R^2$, which also adjusts for the number of dependent features and yields lower values, when features with little entropy towards the target feature are included.
%The $f$-$statistic$ determines how much improvement the independent variables will add to the model compared to being not considered at all.
The $f$-$statistic$ measures the improvement of the model if independent variables were added. 
\paragraph{Logistic Regression for Dichotomous Target.} Logistic regression implies a dichotomous target variable.
The target is described by fitting a logistic function.
Logistic models, as opposed to linear models, do not allow for extracting an $R^2$ quality of fit value.
Therefore, pseudo-$R^2$ values are extracted, such as the \emph{Nagelkerke $R^2$} \cite{Nagelkerke}, which mimics the behavior of the $R^2$.
\emph{Nagelkerke $R^2$} behaves different than $R^2$ values extracted from the linear regression model.
Comparisons have to be handled with care.
\section{Prior and Related Work}
%% https://en.wikipedia.org/wiki/Exploratory_data_analysis#cite_note-Tukey1977-4
Tukey already stated in 1977 that data are too often analyzed solely using confirmatory data analysis \cite{Tukey}.
He emphasized the need to use data to \emph{derive} hypotheses, which can then be tested again.
In this section we present prior and related work trying to achieve this goal.
%The work of M\"{u}hlbacher et al.~\cite{Muehlbacher}, Piringer et al.~\cite{Piringer}, Chan et al.~\cite{Chan} and Zhang et al.~\cite{Zhang2014} are closest related to ours regarding different aspects.
\paragraph{Parameter space analysis using regression models.}
Sedlmair et al.~\cite{Sedlmair} presented a conceptual taxonomy of parameter space analysis.
%It includes a abstract application domain independent data flow model, navigation strategies for exploring parameter spaces as well a characterization of analysis tasks.
%Using the taxonomy, approaches can be classified to find works with similar problems and derive new design ideas and evaluate ideas.
%The taxonomy allows to classify approaches to find works with similar problems to derive new design ideas and evaluate ideas.
Based on their input parameter taxonomy, we are visualizing \emph{model parameters} based on \emph{environmental parameters} in a \emph{global-to-local} navigation strategy.
We solve a \emph{fitting} task by aiming to find models well suited for describing the input data.
%% A Partition-based Framework for Building and Validating Regression Models
The approach of M\"{u}hlbacher et al.~\cite{Muehlbacher} is closest to ours.
They provided a framework for qualitative analyses of relationships and ranking features for numerical target features with regression models.
%Correlations between features with the target yield a ordered matrix-plot, where feature combinations are used to depict models of interest.
%A view for single-relationships with the target and single features as well as a matrix-plot for feature combinations are used to depict models of interest.
%The visualization allows to assess different model complexities.
Existing regression models can be validated and compared using 3D views and 2D slice views.
%We employ the same steps by preprocessing features using correlations with the target and providing an overview over combinations.
M\"{u}hlbacher et al. focused on a smaller number of features, which can be assessed in more detail, yielding a plot matrix view, while we cover more features by abstracting the models.
%We also take categorical targets and regression formula inputs into account.

%% HyperMoVal: Interactive Visual Validation of Regression Models for Real-Time Simulation
Similarly, Piringer et al.~\cite{Piringer} proposed methods for visualizing regression analysis results and properties for developing car engines.
%%
Their main goal is to assess the pairwise influence of independent features w.r.t. the target feature using a plot matrix displaying models as contours.
%%
%They also incorporate 3D visualizations for each pairwise combination, but mainly because of its popularity with the target domain engineers.
%%
Linked views of model deviations allow to select outliers.
%%
This limits the method to comparing a few models at once as the plot matrix gets complex with increasing feature number.
%%
% The main difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%
% We analyze a large amount of simpler models in terms of \emph{different} features.
% %%
% They also focus on metric features, while we process categorical data as well.
%%
Guo et al.~\cite{Guo} presented multi-space visualizations to find linear relationships in the data with focus on extracting groups of best fit.
%%
The data space is visualized using a scatter plot matrix.
%%
Linear models are calculated by defining dependent and independent features.
%%
The model view allows for assessing different models by color coding distances to the line-of-fit.
%%
The model-parameters can then be fine-tuned using line graphs, histograms and model projections.
%%
%% Regression Cube: A Technique for Multidimensional Visual Exploration and Interactive Pattern Finding
%We use the term `Regression Cube' for a comparative 3D representation of multiple regression models.
%%
Chan et al.~\cite{Chan} propose the `Regression Cube', an extension of the 2D scatter plot representation of a linear regression model (incorporating solely metric features) to a 3D Cube.
%%
They group subjects using a set of interaction techniques as well as clustering algorithms to calculate sub-groups, which can then be compared using their cube representation.
%%
Similar to Piringer et al.~\cite{Piringer}, they focus on highlighting details of the included models rather than comparing models consisting of different features.
%%
Insight is derived by subject grouping, which spawns new cube correlations and therefore allows drilling down to the data.
%%
%In contrast to this approach, we focus on comparing models using quality-of-fit measures and do not focus on subdividing the data sets to gain insight.
%%
%We model expert knowledge through the definition of regression formulas, which is not the focus of Chan et al.
%%
Piringer et al.~\cite{Piringer}, Guo et al.~\cite{Guo} and Chan et al.~\cite{Chan} focus on finding and tuning a model for a specific relationship, while we search for models w.r.t. all combinations in the data set.
%%
Instead of analyzing one complex model in detail, we process a large amount of models in terms of \emph{different} features.
%We difference is their focus on analyzing one complex model in detail, yielding extensive plots.
% %%

@@

%% Steenwijk: Integrated Visual Analysis for Heterogeneous Datasets in Cohort Studies
Steenwijk et al.~\cite{Steenwijk} focus on \emph{hypothesis-free} exploration of cohort data.
%%
They employ a framework consisting of feature extraction and visualization to derive dependencies between image- and non-image features.
%%
They incorporate linked views using scatter plots, bar charts, parallel coordinates as well as time plots to display and brush the data.
%%
%Medical image data can be incorporated to investigate potentially interesting individuals.
%%
%%We distinguish to this and other linked-view analysis based approaches by showing abstracted information derived from a preceding analyze-first step to highlight relationships rather than displaying the actual data using various information visualization techniques.
%%
The techniques are useful for selected features or a data set with a small number of features due to their increase in complexity with every additional parameter.
%%
%They are therefore suitable for \emph{small} epidemiological data sets.
%%
Maries et al.~\cite{GRACE} proposed \texttt{GRACE}, a framework for visually exploring correlations of features with spatial and non-spatial geriatric data.
%%
Sparse Partial Least Squares Regression and Tikhonov Regularization are used to produce predictor subsets and quantify correlation strengths.
%%
A multiple view system links spatial with non-spatial data and correlation coefficients are visualized using Kiviat diagrams, correlations are plotted using p-values in the other views.
%%
%Dendograms are used to select different brain regions.
Instead of applying several linked views, we focus on one view of abstracted data.

Generalized Pairs plots (\texttt{GPLOM'S}) extend the concept of scatter plot matrices by the pairwise depiction of heterogeneous data using type-combination-dependent visualizations \cite{GPLOMS, Francois2013}.
%%
Dai et al.~\cite{Dai2005} incorporate a \texttt{GPLOM}-like visualization using choropleth maps mapping spatial data, such as mortality rates together with scatter plots augmented with \emph{Pearsson's} $r$ values.
%%
A \emph{concept map} summarizes features related to a specified disease.
%%
Time-dependent epidemiological data are visualized by Chui et al.~\cite{Chui2011} using multi-panel graphs highlighting risk factor differences with age and gender with regard to influenza- and salmonellosis-associated hospitalizations.
%%
We incorporate the idea of the arrangement of feature combinations as matrix as well as assessing confounding features.
%%

%%
%We also include confounder features (age, gender) into the analysis while focusing on non-time-dependent cohort study data sets.
%%
\paragraph{Statistical analysis.}
%%
%Statistical analysis is an inherent part of Visual Analysis.
%%
%Here we present work showing a strong focus on statistical analysis in this context.
%% Quality Metrics in High-Dimensional Data Visualization - An Overview and Systematization
Bertini et al.~\cite{Bertini} present an overview of quality metrics describing high dimensional-data.
%%
Their refer to their research agenda in our design.
%%
Namely, we apply aspects of perceptual tuning with human pattern recognition of important aspects, scalability between different data set sizes and application testing with domain experts.
%% The Sparse Regression Cube: A Reliable Modeling Technique for Open Cyber-physical Systems
Ahmadi et al.~\cite{Ahmadi} define the \emph{Sparse Regression Cube} that partitions sparse high dimensional data into subspaces, which are then described by their most reliable linear regression model.
%%
%In their application, they try to find most fuel-efficient roads connecting two user-defined landmarks.
%%
They focus on an algebraic representation for efficient regression-model calculation to find the best fit for a subspace.
%%
%The dimensions for the calculations are fixed, while we assess different features for each model calculation and can therefore not benefit from a hierarchical calculation approach.

@@

@@

@@

As described in Section~\ref{sec:RegressionAnalysis}, regression analysis is the statistical tool of choice for analyzing relationships in epidemiological data.
%%
A regression model is based on expert knowledge.
%%
There is no rule how to apply models to a given set of features. Thus, they have to be applied with care.
%%
% \design{
% %%
% Key requirements based on the epidemiological application domain on the 3D Regression Heat Map are:
% %%
% \begin{itemize}
% 	\item Show relationships of features w.r.t. their explanatory power towards the target feature (e.g., a disease or unhealthy condition).
% 	\item Allow for modeling assumptions regarding the target feature.
% 	\item The tool has to be easily accessible and simultaneously secure the data its processing.
% \end{itemize}
% %%
% }
\subsection{Regression Heat Map Description Using Regression Formula Notation}
%%
Expert knowledge modeling is carried out using \emph{regression formulas}.
%Expert knowledge is integrated into the analysis using regression formulas.
%%
The formula input influences the type of the chosen regression method as well as the \emph{independent} features describing the target (recall Sect.~\ref{sec:RegressionAnalysis}).
%%

@@

The number of features calculated by the CFS algorithm is dependent on the information entropy in the data.
%%
In our epidemiological data, we usually observed a number of 10 to 30 features.
%%
The number of selected features using the CFS algorithm reflects their information entropy on the target.
%%
A large list of features is an expression of low correlation to the target feature.
%%
The tradeoff involved using the CFS algorithm is the potential removal of interesting features for the domain expert.
%%
This problem is discussed in the next subsection as part of the 3D representation of the regression results.
%%

@@

\paragraph{3D Regression Heat Map comparison.}
%%
Comparisons were introduced later in the project.
%%
Epidemiologists with focus on statistics pointed out that comparing outcomes of different formulas is suitable for removing the effect of possible confounding features.
%%
\emph{3D Regression Heat Maps} can be compared by creating difference views.
%%
One formula acts as reference.
%%
The absolute difference in the regression metric values with the second formula is calculated. %yielding the difference containing only the differences between the two formulas.
%%
For example, it can be utilized for comparing the influence of a single feature on the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$).

\section{System Design} \label{sec:SystemDesign}
\begin{figure}[htb]
 \centering
 \includegraphics[width=1.0\linewidth]{figures/system}
 \caption{
 %%
 Breast density data set loaded into our prototype.
 %%
 (a) Using the formula input, the user specifies the dependent feature and calculation rules.
 %%
 (b) 3D heat map, showing values above the matrix diagonal as overview.
 %%
 The values of the currently selected slice are mirrored and represented as orange data points on the slicing plane.
 %%
 (c) 2D heat map of the selected slice for feature \emph{Pain/Discomfort}.
 %%
 }
  \label{fig:System}
\end{figure}
%%
%Albeit our focus on epidemiological data sets, we designed our system to be open with open formats as input interfaces .
%%
\noindent We designed our system to be openly accessible and easy to use.
%%
With open formats as input interfaces, the application can be extended to non-epidemiological data sets.
%%
The focus lies on creating an overview visualization and gaining insight into relationships of the data, which triggers further analyses with other (statistical) tools.
%%
This is, however, out of the scope of this work.
%%
Therefore, the system has to be intuitive and comprehensive in order to be adapted by domain experts.
